Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-12 WSSSPE3/report/wg_appendix_best_practices.tex
more comments/edits

\subsection{Group Members}

@@

\subsection{Summary of Discussion}
Core questions that will need to be explored are in knowledge management, 
(transitions between people), reliability (reproducibility), usability and how a software tool becomes part of the core workflow of well identified users (stakeholders)
relating to tool success and hence sustainability. Ideas 
that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment e.g.(new hardware, new methodology) ;
\item Dimensions of sustainability - economic, technical, environmental, 
declining interest in primary application area), social.
\end{itemize}

Sustainability requires community participation in code development and/or a wide adoption of software. The larger the community base is using a software, the better are the funding possibilities and thus also the sustainability options. Additionally developer commitment to an application is essential and experience shows that especially software packages with an evangelist imposing strong inspiration and discipline is required to achieve sustainability.
While a single person can push sustainability to a certain level, open source software also needs sustained commitment from the developer community. Such sustained commitments include diverse tasks and roles, which can be fulfilled by diverse developers with different knowledge levels. Besides developing software and appropriate software management with measures for extensibility and scalability of the software, active (expertise) support for users via a user forum with a quick turnaround is crucial. The barrier to entry the community as user as well as developer has to be as low as possible.

@@

The opportunity lays in the collaboration on a white paper, which will be revisited regularly for further improvements and enhances the state on best practices resulting in a peer-reviewed paper. This way we would like to reach a wide community. But these are also the challenges and obstacles - to get everyone to write on the paper and reach the community.

@@

The key next steps are to write the introduction, reach out to the co-authors and agree on a scope of the white paper.

>>>>

\subsection{Group Members}
\katznote{add affils for all, please}

@@

\subsection{Summary of Discussion}

Core questions that will need to be explored are in knowledge management, 
(transitions between people), reliability (reproducibility), usability, and how a software tool becomes part of the core workflow of well identified users (stakeholders)
relating to tool success and hence sustainability. \katznote{prev sentence is complex and awkward} Ideas 
that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment e.g.(new hardware, new methodology) ;
\item Dimensions of sustainability -- economic, technical, environmental, 
declining interest in primary application area), \katznote{not sure what the prev. comment goes with} social.
\end{itemize}

Sustainability requires community participation in code development and/or a wide adoption of software.
The larger the community base is using a piece of software, the better are the funding possibilities and thus also the sustainability options.
Additionally developer commitment to an application is essential and experience shows that software packages with an evangelist imposing strong inspiration and discipline are more likely to achieve sustainability.
While a single person can push sustainability to a certain level, open source software also needs sustained commitment from the developer community.
Such sustained commitments include diverse tasks and roles, which can be fulfilled by diverse developers with different knowledge levels.
Besides developing software and appropriate software management with measures for extensibility and scalability of the software, active (expertise) support for users via a user forum with a quick turnaround is crucial.
The barrier to entry for the community as users as well as developers has to be as low as possible.

@@

The opportunity lies in collaboration on a white paper, which will be revisited regularly for further improvements, to enhance knowledge of the state of best practices, resulting in a peer-reviewed paper. We would like to reach a wide community by doing this. But these are also the challenges and obstacles -- to get everyone to contribute to the paper and to reach the community.

@@

The key next steps are to write the introduction, reach out to the co-authors, and to agree on the scope of the white paper.
